---
title: "Covid Assigment"
date: 2020-11-28
author: "Yuri Almeida Cunha"
output: pdf_document
        #html_document
  
---

```{r, echo= FALSE , include = FALSE }
# Clear memory
rm(list=ls())

# Packages to use
library(tidyverse)
# For scaling ggplots
require(scales)
# Estimate piecewise linear splines
#install.packages("lspline")
library(lspline)
# Estimate robust SE
#install.packages("estimatr")
library(lmtest)
#check Homoscedastics 
library(estimatr)
# Compare models with robust SE
#install.packages("texreg")
library(texreg)
# For different themes
#install.packages(ggthemes)
library(ggthemes)
#Linear Hypothesis
library(car)
# For display of data tables
#install.packages(knitr)
library(knitr)
#install.packages(pander)
library(pander)

# Call the data from local file

data_path <- 'https://raw.githubusercontent.com/yurialmeida5/Covid_Assigment_Yuri_Cunha/main/data/clean/covid_pop_09_29_2020_clean.csv'

df <- read_csv(data_path)

```


# Executive Summary

On the following project, I checked the possible association between the number of confirmed Covid cases and their number of deaths associated with the disease by country on a given day. Using the Weighted - OLS model, considering the population size as the weight, I was able to reach a well-fitted model for the following dataset. The models states that the increase of the number of confirmed cases increases on average the number of deaths associated with the disease, and that the population numbers play an important role on that. It's important to mention that the data acquired may have some integrity and reliability issues due some government data collection standards. 

# Research Question and Variable Description

Research Question:   
Is there any association between number of deaths and confirmed cases on this dataset? If yes, what is it?  
Variables Description:  
death (Explanatory): number of accumulated covid deaths until a given day by country    
confirmed (Dependent): number of accumulated confirmed covid cases until a given day by country 
Sample: number of cases and deaths (until 29/09)  / Population: Total number of cases and deaths whole Covid season
Possible Quality Issues: Reliability (Data integrity - Governmental Interests) , Content (Caused by Covid or simply with Covid at death moment?

# Scaling and Dropping Values

No scaling was used, used actual absolute numbers.    
Few observations with really lower deaths and confirmed cases.    
Removed all countries with low confirmed cases and really low population - not relevant for analysis.   
Virus not sufficiently spread, and also helps to eliminate countries with specific geographic characteristics.  
Ex: Brunei, Comoros, Dominica, Fiji and some other islands.

```{r, echo = FALSE}

df <- df %>% filter(confirmed > 5000,
                            population > 100000)
```

# Variable Check 

```{r, echo= FALSE}


summary_x <- df %>%
  summarise(
    variable = 'confirmed',
    mean = mean(confirmed),
    median = median(confirmed),
    min = min(confirmed),
    max = max(confirmed),
    sd = sd(confirmed)
  )
summary_y <- df %>%
  summarise(
    variable = 'death',
    mean = mean(death),
    median = median(death),
    min = min(death),
    max = max(death),
    sd = sd(death)
  )

table_summary <- add_row(summary_x,summary_y)
kable(table_summary)


```


```{r, echo = FALSE, fig.show='hold',figures-side, out.width='50%'}

hist(df$death, main = "Covid Deaths (until 09-29-20)", xlab="Number of Deaths", 
     ylab = "Frequency", border="blue", col="lightgreen", las=1, 
     breaks=10)

hist(df$confirmed, main = "Covid Confirmerd Cases (until 09-29-20)", xlab="Number of Confirmed Cases", 
     ylab = "Frequency", border="blue", col="lightgreen", las=1, 
     breaks=10)

```

Deaths and Confirmed Cases:
Right tail Distributions: median < mean, outliers distributed on the right part of the graph


# Log Transformations

Comparing the p-values (significance) and the R^2 adjusted (highness) for the four types of possible transformations (check the appendix),it´s possible to check that the lvl_lvl and log_log represented the best possible models to choose.  
The models lvl_lvl and log_log showed good p and R^2-Ajusted values,checking the graphs its possible to determine that taking logs will be needed for both cases.Both deaths and confirmed cases are highly right-tailed skewed distributions with a few outliers.  
Although, level is easier to interpret in general, the logs in this case produce a better fit, 
making the association pattern between them better to identify and analyze.   
Absolute values are not that relevant in both cases.  

```{r, echo = FALSE}

df <- df %>% mutate(ln_death = log(death),
                    ln_confirmed = log(confirmed))

```

# Models

Based on model comparison our chosen model is reg4 - Weighted-OLS   
## Substantive:   
- The model demonstrated a really great fit for the dataset, and can be used for prediction   
- The beta parameter states that for the following dataset an increase of 10% on the number of confirmed cases will increase, on average, the number of deaths in 9.4%  
- The alfa in that case represents the mean effect on number of deaths of variables not included in the model.  
## Statistical: 
- The models showed really similar statistical results (p-values and r-adjusted), capturing the variation really well.   
- As logically expected, the weighted one by population size could demonstrate a even better fit.  

```{r, echo = FALSE}

reg4 <- lm_robust(ln_death ~ ln_confirmed, data = df , weights = population)

```


# Hypothesis Testing

```{r, echo= FALSE}

summary <- summary(reg4)
kable(summary$coefficients)

```

Beta coefficient test is equal to 0?
Test: H0:  = 0, HA: ln_confirmed <> 0 
Can reject null hypothesis: equal to 0 

# Residual Analysis 

```{r, echo = FALSE}

# Get the predicted y values from the model
df$reg4_y_pred <- reg4$fitted.values
# Calculate the errors of the model
df$reg4_res <- df$ln_death - df$reg4_y_pred 

```

## Top 5 Countries with lower death than predicted by regression

```{r, echo = FALSE}

top5_lower <- df %>% top_n( -5 , reg4_res ) %>% 
  select( country , ln_death , reg4_y_pred , reg4_res )

pander(top5_lower)

```

## Top 5 Countries with higher death than predicted by regression

```{r, echo = FALSE}

top5_upper <- df %>% top_n( 5 , reg4_res ) %>% 
  select( country , ln_death , reg4_y_pred , reg4_res )

pander(top5_upper)


```


\newpage

# Appendix

## Log Transformations Decision

### Check homoskedastic - Breusch-Pagan Test

```{r}

lvl_lvl <- lm( death ~ confirmed , data = df)
log_lvl <- lm( death ~ ln_confirmed , data = df )
lvl_log <- lm( ln_death ~ confirmed , data = df )
log_log <- lm( ln_death ~ ln_confirmed , data = df )

bptest( lvl_lvl ) # small p-value can´t reject the null hypothesis (homoscedasticity)
bptest( log_lvl ) # small p-value can´t reject the null hypothesis (homoscedasticity)
bptest( lvl_log ) # big p-value can t reject the null hypothesis (homoscedasticity)  
bptest( log_log ) # big p-value can reject the null hypothesis (homoscedasticity)

# Use lm_robust lvl_log / log_log
lvl_log <- lm_robust( ln_death ~ confirmed , data = df , se_type = "HC2" )
log_log <- lm_robust( ln_death ~ ln_confirmed , data = df , se_type = "HC2" )

```

### Graphical Check

```{r}

#Death - confirmed : level-level 

ggplot( df , aes(x = confirmed , y = death)) +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = " Number of Confirmed Cases", y = "Number of Deaths") 

# Death - confirmed : log-log
ggplot( df , aes(x = confirmed, y = death ))  +
  geom_point() +
  geom_smooth(method="loess")+
  labs(x = " Number of Confirmed Cases , ln scale )",y = "Number of Deaths, ln scale)") +
  scale_x_continuous( trans = log_trans(),  breaks = c(10000,50000,1000000))+
  scale_y_continuous( trans = log_trans())

 

```

Considering the graphs you can definetily see that you reach a better fit using the log_log, instead of the lvl_lvl. The distribution of deaths is a really right-tailed skewed one.

## Model Comparison - Graphs

```{r, echo = FALSE}

## Estimating Different Models

reg1 <- lm_robust( ln_death ~ ln_confirmed , data = df , se_type = "HC2" )
summary( reg1 )

# Visual inspection:
ggplot( data = df, aes( x = ln_death, y = ln_confirmed ) ) + 
  geom_point( color='blue') +
  geom_smooth( method = lm , color = 'red' )


### Quadratic Linear Regression

### Add quadratic variable to the dataframe

df <- df %>% mutate(ln_confirmed_sq = ln_confirmed^2)

# Second and third model with gdptot
reg2 <- lm_robust( ln_death ~ ln_confirmed + ln_confirmed_sq , data = df )
summary( reg2 )

# Visual inspection:
ggplot( data = df, aes( x = ln_confirmed , y = ln_death ) ) + 
  geom_point( color='blue') +
  geom_smooth( formula = y ~ poly(x,2) , method = lm , color = 'red' )

# Piecewise linear spline:

cutoff <- 400000
cutoff_ln <- log( cutoff )

reg3 <- lm_robust(ln_death ~ lspline( ln_confirmed , cutoff_ln ), data = df )
summary( reg3 )

ggplot( data = df, aes( x = ln_confirmed, y = ln_death ) ) + 
  geom_point( color='blue') +
  geom_smooth( formula = y ~ lspline(x,cutoff_ln) , method = lm , color = 'red' )

# Weighted-OLS

reg4 <- lm_robust(ln_death ~ ln_confirmed, data = df , weights = population)
summary( reg4 )

ggplot(data = df, aes(x = ln_confirmed, y = ln_death)) +
  geom_point(data = df, aes(size=population),  color = 'blue', shape = 16, alpha = 0.6,  show.legend=F) +
  geom_smooth(aes(weight = population), method = "lm", color='red')+
  scale_size(range = c(1, 15)) +
  labs(x = "ln(Number of Confirmed Cases) ",y = "ln(Number of Deaths)")
  

```

## Model Comparison - HTML

```{r}

data_out <- 'D:/CEU/Data_Analysis_2/Covid_Assigment_Yuri_Cunha/out/'
htmlreg( list(reg1 , reg2 , reg3 , reg4),
         type = 'html',
         custom.model.names = c("Linear Regression", "Quadratic (Linear) Regression","Piecewise Linear Spline",
                                "Weighted-OLS"),
         caption = "Modelling number of deaths by covid based on number of confirmed cases",
         file = paste0(data_out,'model_comparison.html'), include.ci = FALSE)


```

For the analysis, check the main body 







